---
title: "DataAnalysis"
author: "Nikola_Stankovic"
---
##Data Analysis
Import the dataset:
``````{r}
fdata <- read.csv("http://archive.ics.uci.edu/ml/machine-learning-databases/forest-fires/forestfires.csv")
#fdata <- read.csv("C:/CKME136/Forest Fires/forestfires.csv")
str(fdata)
nrow(fdata)
ncol(fdata)
names(fdata)
```
Check for missing values:
```{r}
(lapply (fdata, function (x) which (is.na (x))))
```

Almost 50% of the area data has zeroes. As per the article these are valid as they represent small fires < 100 m2:
```{r}
zeroes <- (subset(fdata, fdata$area==0))
(nrow(zeroes) / nrow(fdata))
```

Data is very skewed to the right as per histogram:
```{r}

hist((fdata$area),40, ylim = c(0, 500), xlab = 'Area', main = 'Histogram of Area')
mtext(c("Figure 1"),side=1,line=4,at=c(0,0), cex = 0.7)
```

Transform area to log function to improve linear relationship with predictors
```{r}
fdata <- within(fdata, {arealog <- log(area + 1)})
hist((fdata$arealog),40, ylim = c(0, 500), xlab = 'Log of Area',  main = 'Histogram of Log of Area')
mtext(c("Figure 2"),side=1,line=4,at=c(0,0), cex = 0.7)
```

Correlation plot with actual numeric correlation
```{r}
panel.pearson <- function(x, y, ...) {
  horizontal <- (par("usr")[1] + par("usr")[2]) / 2
  vertical <- (par("usr")[3] + par("usr")[4]) / 2
  text(horizontal, vertical, format((cor(x,y)), digits=2))
}
```


```{r}
pairs(fdata[,5:13], lower.panel=panel.pearson, main = "Scatterplots-Pearson Correlation")
mtext(c("Figure 3"),side=1,line=4,at=c(0,0), cex = 0.7)
```

Both Area and Log of Area have very low correlation with all the predictor variables as per scatter plot above

Scatter plots for each numeric variable individually:
```{r}
par(mfrow=c(3,3))
plot(arealog ~ FFMC + DMC + DC + ISI + temp + RH + wind + rain,
     data = fdata)
par(mfrow=c(1,1))
```
Plotting packages:
```{r}
install.packages('ggplot2',repos = "http://cran.rstudio.com/")
library(ggplot2)
library(grid)
library(gridExtra)
```

Area per month box plot
```{r}
grid.arrange(qplot(x = month, xlab = "Month", y = area, ylab =  "Area"
, data = fdata, geom = 'boxplot'), bottom = textGrob("Figure 4", x = 1, hjust = 10, gp = gpar(fontsize = 9)))
```

There are a couple of outliers in the box plot above, but I am keeping them as that is real data 

Group months into seasons:
```{r}
fdata$season <- rep("spring", nrow(fdata))
for (i in 1:nrow(fdata)){
  if (fdata$month[i] %in% c("feb","jan","dec")) fdata$season[i] <- "winter"
  if (fdata$month[i] %in% c("oct","nov","sep")) fdata$season[i] <- "autumn"
  if (fdata$month[i] %in% c("aug","jul","jun")) fdata$season[i] <- "summer"
}
fdata$season <- as.factor(fdata$season)
```

Area per season box plot:
```{r warning=FALSE}
qplot(x = season, y = area , data = fdata, geom = 'boxplot', coord_cartesian(ylim = c(-2,8))) 
```

Factor days in weekdays vs weekend classes
```{r}
fdata$wday <- rep("weekend", nrow(fdata))
for (i in 1:nrow(fdata)){
  if (fdata$day[i] %in% c("mon","tue","wed","thu","fri")) fdata$wday[i] <- "weekday"
}
fdata$wday <- as.factor(fdata$wday)
summary(fdata$wday)
```

Convert 9 X 9 grid to 3 X 3 grid
```{r}
fdata$X <- as.character(fdata$X)
fdata$Y <- as.character(fdata$Y)
fdata$XY <- with(fdata, paste(X, Y))
for (i in 1:nrow(fdata)){
  if (fdata$XY[i] %in% c("1 1","2 1","3 1","1 2","2 2","3 2","1 3","2 3","3 3")) fdata$XY[i] <- "one"
  if (fdata$XY[i] %in% c("1 4","2 4","3 4","1 5","2 5","3 5","1 6","2 6","3 6")) fdata$XY[i] <- "two"
  if (fdata$XY[i] %in% c("1 7","2 7","3 7","1 8","2 8","3 8","1 9","2 9","3 9")) fdata$XY[i] <- "three"
  if (fdata$XY[i] %in% c("4 1","5 1","6 1","4 2","5 2","6 2","4 3","5 3","6 3")) fdata$XY[i] <- "four"
  if (fdata$XY[i] %in% c("4 4","5 4","6 4","4 5","5 5","6 5","4 6","5 6","6 6")) fdata$XY[i] <- "five"
  if (fdata$XY[i] %in% c("4 7","5 7","6 7","4 8","5 8","6 8","4 9","5 9","6 9")) fdata$XY[i] <- "six"
  if (fdata$XY[i] %in% c("7 1","8 1","9 1","7 2","8 2","9 2","7 3","8 3","9 3")) fdata$XY[i] <- "seven"
  if (fdata$XY[i] %in% c("7 4","8 4","9 4","7 5","8 5","9 5","7 6","8 6","9 6")) fdata$XY[i] <- "eight"
  if (fdata$XY[i] %in% c("7 7","8 7","9 7","7 8","8 8","9 8","7 9","8 9","9 9")) fdata$XY[i] <- "nine"
}
fdata$XY <- as.factor(fdata$XY)
str(fdata)
```

Test for multicollinearity; 
VIF value greater than 5 indicates multicollinearity between predictor

```{r}
fdata.vif <- fdata[c(5:12)]
install.packages('usdm',repos = "http://cran.rstudio.com/")
library(usdm)
(vif_output <- vif(fdata.vif))
write.table(vif_output, "C:/CKME136/Forest Fires/Vifdata.txt", sep="\t", row.names = F)
```

The highest VIF value is 2.66, means no multicollinearity. No need to remove any numeric variables

SAVE DATASET IN A TEMPORARY DATA FRAME FOR  THE LATER ANALYSIS
```{r}
fdatatemp <- fdata
save(fdatatemp,file="C:/CKME136/Forest Fires/fdata.RDa")
```