---
title: "LinearModels"
author: "Nikola_Stankovic"
---

###Import prepared data set from the Data Analysis step
```{r}
load("C:/CKME136/Forest Fires/fdata.RDa")
fdata <- fdatatemp
```

Package to perform stepwise linear regression:
```{r}
install.packages("MASS",repos = "http://cran.rstudio.com/")
library(MASS)
```

Package used to perform cross-validation calls to 'cross-validation-lm.R' script
```{r}
install.packages("cvTools",repos = "http://cran.rstudio.com/")
library(cvTools)
```

Exclude original spatial and 
temporal predictors and keep the transformed ones: season, wday and XY (simplified coordinates)
```{r}
fdata <- fdata[,c(5:17)]
```


###1.Apply linear model to using weather and fire index numeric predictors
```{r}
lm.wf <- lm(formula = arealog ~ FFMC + DMC + DC + ISI + temp + RH + rain + wind,data= fdata)
summary(lm.wf)
```
The model has very low R-squared and adjusted R-squared
Multiple R-squared:  0.01988
Adjusted R-squared:  0.004446 

Test the model using 10-fold cross-validation repeated 30 times:
```{r}
sel_var_lm <- "FFMC + DMC + DC + ISI + temp + RH + rain + wind" 
source('C:/CKME136/Forest Fires/FinalSubmission/cross-validation-lm.R', echo = TRUE)
```
##### > RMSE
##### [1] 64.44124
##### > MAD
##### [1] 12.93817

###1a.Stepwise selection for weather and fire numeric variables
```{r  results="hide"}
stepB <- stepAIC(lm.wf, scope=list(upper=~.,lower=NULL, direction= "backward", trace=TRUE))
```

```{r}
summary(stepB)
```
The suggested best model includes FFMC, RH, wind

Multiple R-squared:  0.01425
Adjusted R-squared:  0.008485

Test the model:
```{r}
sel_var_lm <- "DMC + RH + wind" 
source('C:/CKME136/Forest Fires/FinalSubmission/cross-validation-lm.R', echo = TRUE)
```
##### > RMSE
##### [1] 64.44783
##### > MAD
##### [1] 12.95806

###2.Linear model all the predictors (spatial and temporal are categorized in less number of classes)
```{r}
lm.all <- lm(formula = arealog ~ .-area, data= fdata)
summary(lm.all)
```
Multiple R-squared:  0.05259
Adjusted R-squared:  0.01835 

Test the model
```{r}
sel_var_lm <- ".-area" 
source('C:/CKME136/Forest Fires/FinalSubmission/cross-validation-lm.R', echo = TRUE)
```
##### > RMSE
##### [1] 64.33444
##### > MAD
##### [1] 12.85998

###2a. Apply stepwise for on the previous model
```{r results="hide"}
stepC <- stepAIC(lm.all, scope=list(upper=~.,lower=NULL, direction= "backward", trace=TRUE))
```
```{r}
summary(stepC)
```
Suggested predictors: DMC + wind + temp + season

Multiple R-squared:  0.0377
Adjusted R-squared:  0.02638 

Test the model:
```{r}
sel_var_lm <- "DMC + temp + wind + season" 
source('C:/CKME136/Forest Fires/FinalSubmission/cross-validation-lm.R', echo = TRUE)
```
#####> RMSE
#####[1] 64.39118
#####> MAD
#####[1] 12.89213

###3. Polynomial regression
Test if quadratic model is better than regular linear for each predictor separately
```{r}
for (i in 1:4) {
  lm.ffmc <- lm(formula = arealog ~  poly(fdata$FFMC, i),data= fdata)
  print(summary(lm.ffmc))
}
for (i in 1:4) {
  
  lm.dmc <- lm(formula = arealog ~  poly(fdata$DMC, i),data= fdata)
  print(summary(lm.dmc))
}

for (i in 1:4) {
  lm.isi <- lm(formula = arealog ~  poly(fdata$ISI, i),data= fdata)
  print(summary(lm.isi))
}
for (i in 1:4) {
  lm.dc <- lm(formula = arealog ~  poly(fdata$DC, i),data= fdata)
  print(summary(lm.dc))
}
for (i in 1:4) {
  lm.temp <- lm(formula = arealog ~  poly(fdata$temp, i),data= fdata)
  print(summary(lm.temp))
}

for (i in 1:4) {
  lm.rh <- lm(formula = arealog ~  poly(fdata$RH, i),data= fdata)
  print(summary(lm.rh))
}

for (i in 1:4) {
  lm.wind <- lm(formula = arealog ~  poly(fdata$wind, i),data= fdata)
  print(summary(lm.wind))
}

for (i in 1:4) {
  lm.rain <- lm(formula = arealog ~  poly(fdata$rain, i),data= fdata)
  print(summary(lm.rain))
}
```

3 out of 4 weather predictors (wind, temperature and rain) showed better results(smaller p-value) when polynomials applied
I will take the best model so far (stepC) and add quadratic terms wind, temp and also rain
```{r}
lm.poly <- lm(formula = arealog ~  DMC + poly(temp,3) + poly(wind,3) + poly(rain,3) + 
                season, data = fdata)
summary(lm.poly)
```
Multiple R-squared:  0.06256
Adjusted R-squared:  0.03834 

Test the model:
```{r}
sel_var_lm <- "DMC + poly(temp,3) + poly(wind,3) + poly(rain,3) + 
                season" 
source('C:/CKME136/Forest Fires/FinalSubmission/cross-validation-lm.R', echo = TRUE)
```
#####> RMSE
#####[1] 64.34547
#####> MAD
#####[1] 12.78232

### 3a Add quadratic term for DC
```{r}
lm.polyDC <- lm(formula = arealog ~  DMC + poly(temp,3) + poly(wind,3) + poly(rain,3) + poly(DC,2)  +
                season, data = fdata)
summary(lm.polyDC)
```

Multiple R-squared:  0.06999	
Adjusted R-squared:  0.04215 

Test the model:
```{r}
sel_var_lm <- "DMC + poly(temp,3) + poly(wind,3) + poly(rain,3) + poly(DC,2)  +
                season" 
source('C:/CKME136/Forest Fires/FinalSubmission/cross-validation-lm.R', echo = TRUE, max.deparse.length=1e3)
```
#####> RMSE
#####[1] 64.30157
#####> MAD
#####[1] 12.75959

### 4. Use the model created in step 2a and add interactions to it
```{r results="hide"}
lm.ie <- lm(formula = arealog ~ (DMC + wind + temp + season)^2, data= fdata)
stepD <- stepAIC(lm.ie, scope=list(upper=~.,lower=NULL, direction= "backward", trace=TRUE))
```
```{r}
summary(stepD)
```
Multiple R-squared:  0.05421
Adjusted R-squared:  0.03552


Test the model
```{r}
sel_var_lm <- "DMC + wind + temp + season + DMC:temp + DMC:season"
source('C:/CKME136/Forest Fires/FinalSubmission/cross-validation-lm.R', echo = TRUE)
```
#####> RMSE
#####[1] 64.35089
#####> MAD
#####[1] 12.79311


### 4a Tests interactions for all predictors
```{r}
lm.ieall <- lm(formula = arealog ~ (.-area)^2, data= fdata)
```

```{r results="hide"}
stepE <- stepAIC(lm.ieall, scope=list(upper=~.,lower=NULL, direction= "backward", trace=TRUE))
```

```{r}
summary(stepE)
```

The best model has improved significantly: 
Multiple R-squared:  0.1533
Adjusted R-squared:  0.07242


Test the model
```{r}
sel_var_lm <- "DMC + DC + ISI + temp + RH + wind + rain + 
    season + XY + DMC:DC + DMC:RH + DMC:season + DC:XY + ISI:temp + 
ISI:season + ISI:XY + temp:rain + temp:season + RH:season + 
ISI:rain"
source('C:/CKME136/Forest Fires/FinalSubmission/cross-validation-lm.R', echo = TRUE)
```
#####> RMSE
#####[1] 64.06452
#####> MAD
#####[1] 12.47535
