---
title: "LogisticRegression"
author: "Nikola_Stankovic"
---

# Logistic Regression

install package used for logistic regression
```{r}
install.packages("ROCR",repos = "http://cran.rstudio.com/")
library(ROCR)
install.packages("gplots",repos = "http://cran.rstudio.com/") 
library(gplots)
```
```
Package used to perform cross-validation calls to 'cross-validation-rt.R' script
```{r results="hide"}
install.packages("cvTools",repos = "http://cran.rstudio.com/")
library(cvTools)
```

Keep actual area as the response, remove log of area
```{r}
load("C:/CKME136/Forest Fires/fdata.RDa")
fdata <- fdatatemp
fdata <- fdata[,c(5:17)]
fdata_lr <- fdata[,c(-10)]
```

create new factor response variable based on area size 
```{r}
fdata_lr$classarea <- as.factor(ifelse(fdata_lr$area == 0, 0, 1))
```
remove area 
```{r}
fdata_lr <- fdata_lr[,c(-9)]
```
###1. Stepwise regression to select the best predictors
```{r}
full_model <- glm(classarea ~ RH + wind + temp + FFMC + DMC + DC + ISI + season + wday + XY + rain,family = "binomial", data=fdata_lr, maxit=100)
no_model <- glm(classarea ~ 1, family = "binomial", data=fdata_lr)
stepwise_model <- step(no_model, list(lower=formula(no_model), upper=formula(full_model)), direction="both", trace = 0)
summary(stepwise_model)
```
selected model predictors are DC and wind only

perform ROC test for the entire dataset 
```{r}
p <- predict(stepwise_model)
pr <- prediction(p, fdata_lr$classarea)
prf <- performance(pr, measure = "tpr", x.measure = "fpr")
plot(prf)
auc <- performance(pr, measure = "auc")
auc <- auc@y.values[[1]]
auc
```
AUC value of 56% is considered a bad predictor model 

Test the model using 10-fold cross-validation repeated 30 times to get more accurate results
```{r}
sel_var_lr <- "DC + wind" 
source('C:/CKME136/Forest Fires/FinalSubmission/cross-validation-lr.R', echo = TRUE)
```
AUC value after cross-validation is 56%

###2. All predictors except rain
```{r}
sel_var_lr <- "RH + wind + temp + FFMC + DMC + DC + ISI + season + wday + XY" 
source('C:/CKME136/Forest Fires/FinalSubmission/cross-validation-lr.R', echo = TRUE)
```
AUC value of 63% - an improvement over the previous model

###3. Model with interactins and poynomials 
```{r}
sel_var_lr <- "DMC + poly(temp,3) + poly(wind,3) + poly(rain,3) + poly(DC,3)  +
                season + XY + DMC:DC + DMC:RH + DMC:season  +  temp:season + RH:season" 
source('C:/CKME136/Forest Fires/FinalSubmission/cross-validation-lr.R', echo = TRUE)
```
AUC value of 69.9% - an improvement over the previous model

Visual representation of ROC for the best model using entire dataset
```{r}
lr.best.model <- glm(classarea ~ DMC + poly(temp,3) + poly(wind,3) + poly(rain,3) + poly(DC,3)  +
                       season + XY + DMC:DC + DMC:RH + DMC:season  +  temp:season + RH:season
                     ,family = "binomial", data=fdata_lr, maxit=100)
p <- predict(lr.best.model)
pr <- prediction(p, fdata_lr$classarea)
prf <- performance(pr, measure = "tpr", x.measure = "fpr")
plot(prf, main = "ROC for the logistic model using predictors: DMC + poly(temp,3) + poly(wind,3) + poly(rain,3) + 
poly(DC,3)  +  season + XY + DMC:DC + DMC:RH + DMC:season  +  temp:season + RH:season", cex.main = .65 )
```







