---
title: "RegressionTree-RandomForest"
author: "Nikola_Stankovic"
---

## Regression Tree 

Packages required to build and plot regression trees
```{r results="hide"}
install.packages("rpart",repos = "http://cran.rstudio.com/")
library(rpart)
install.packages("rpart.plot",repos = "http://cran.rstudio.com/")
library(rpart.plot)
```
Package used to perform cross-validation calls to 'cross-validation-rt.R' script
```{r results="hide"}
install.packages("cvTools",repos = "http://cran.rstudio.com/")
library(cvTools)
```

###Import prepared data set from the Data Analysis step
```{r}
load("C:/CKME136/Forest Fires/fdata.RDa")
fdata <- fdatatemp
```

Keep log transformation of area (arealog) as the dependent variable, remove area
```{r}
fdata <- fdata[,c(5:17)]
fdata_rt <- fdata[,c(-9)]
```

###1. Weather variable combination 
I tried different number of minbuckets, and the minimum  number of buckets = 8  produced the same results in the
author's paper
```{r}
sel_var <- "temp + RH + rain + wind"
nofbuckets <- 8 #default value for minbuckets
cp_value <- 0 # used for pruning
```

Validate tree using cross validation
```{r}
source('C:/CKME136/Forest Fires/FinalSubmission/cross-validation-arealog-rt.R', echo = TRUE)
```
#####> RMSE
#####[1] 64.49935
#####> MAD
#####[1] 13.15522

The cross validation results using 8 as the minimum number of observations in a terminal 
leaf are in line with the paper (RMSE 64.4 and MAD 13.18)


Visual representation of the tree using entire dataset:
```{r}
rt.w <- (rpart(as.formula(paste("arealog ~", sel_var)), method = "anova", 
                   data = fdata_rt, control = rpart.control(minbucket = nofbuckets)))
rpart.plot(rt.w, type =1, digit=3, fallen.leaves = TRUE)
summary(rt.w)
```
Variable importance
temp 66  wind 18  RH 17 
       
```{r}
printcp(rt.w)
```
The CP table above shows that nsplit = 0 is the best model
This means the average of the entire dataset will always be used for any future predicitons 
This is not a good prediction 


The code below is shown how prunining is doNe in rpart package, however I will not use the prunned tree for any reference
```{r results="hide"}
bestcp <- rt.w$cptable[which.min(rt.w$cptable[,"xerror"]),"CP"]
rt.w.pruned <- prune(rt.w, cp = bestcp)
rpart.plot(rt.w.pruned, type =1, digit=3, fallen.leaves = TRUE)
```

###2 decision tree model using all predictors with minbuckets of 29
```{r}
sel_var <- "."
nofbuckets <- 29 # 
cp_value <- 0
```

visual representation of the tree
```{r}
rt.all <- (rpart(as.formula(paste("arealog ~", sel_var)), method = "anova", 
                    data = fdata_rt, control = rpart.control(minbucket = nofbuckets)))
rpart.plot(rt.all, type =1, digit=3, fallen.leaves = TRUE)
```

Validate the tree using cross validation

```{r}
source('C:/CKME136/Forest Fires/FinalSubmission/cross-validation-arealog-rt.R', echo = TRUE)
```
#####> RMSE
#####[1] 64.391
#####> MAD
#####[1] 13.08937

```{r}
printcp(rt.all)
```

## Random Forest

 Random Forest regression package
```{r}
install.packages("randomForest",repos = "http://cran.rstudio.com/")
library(randomForest)
# Package used to perform cross-validation calls to 'cross-validation-rt.R' script
install.packages("cvTools",repos = "http://cran.rstudio.com/")
library(cvTools)
```

Keep log transformation of area (arealog) as the dependent variable, remove area:
```{r}
fdata <- fdatatemp
fdata <- fdata[,c(5:17)]
fdata_rf <- fdata[,c(-9)]
```
1 - Weather model
```{r}
sel_var_rf <- "temp + RH + rain + wind"
nooftrees <- 100
```

Execute cross validation
```{r}
source('C:/CKME136/Forest Fires/FinalSubmission/cross-validation-arealog-rf.R', echo = TRUE)
```

100 trees - thsi is in line with the paper (RMSE 64.4 and MAD 12.93)

#####> RMSE
#####[1] 64.43702
#####> MAD
#####[1] 12.93335


2 - all predictors
```{r}
sel_var_rf <- "."
nooftrees <- 100
```

Execute cross validation

```{r}
source('C:/CKME136/Forest Fires/FinalSubmission/cross-validation-arealog-rf.R', echo = TRUE) 
```

#####> RMSE
#####[1] 64.38877
#####> MAD
#####[1] 13.16288



